---
title: "R Recommendation engine"
output: html_notebook
---

#Recommendation engine concept

Source: https://www.data-mania.com/blog/how-to-build-a-recommendation-engine-in-r/


Is an automated decision engine that evaluates similarities between people and/or items in order to make recommendations about what items go well together.Most common application is ecommerce: buyers are satisfied with an ideal combination of products and sellers are happy because more sales and a higher ROI.

Collaborative filtering algorithms uses data from users in the same platform to make predictions about items that have strong history of being purchased together. There are two types: User-based collaborative (depending or similar profiles) or item-based collaborative (similarities between items).

advantages: niche items could be suggested, products from different category could be suggested, only based on rating on items could suggest items purchased by similar users.

Not only looking at how many times an item was bought, but recommending rated high by similar users. Real-life recommenders hybridize the techniques and could get complex using solid reasoning to get best logic decisions.

ML algorithms: clustering. Depends of project objectives.

Another type is content-based recommenders when there is not much data available.  


data: https://grouplens.org/datasets/movielens/
demo: http://files.grouplens.org/datasets/movielens/ml-latest-small.zip

#coding 

```{r}
##set working directory

setwd("C:/Users/Moanesga/Desktop/Projects/Recommendation engine/ml-latest-small/")

movies = read.csv(file = "movies.csv")
links = read.csv("links.csv")
ratings = read.csv("ratings.csv")
tags = read.csv("tags.csv")

#head(tags)

```

#Import the reshape2 library. 
#Use the file install.packages("reshape2") if the package is not already installed

```{r}
#install.packages("reshape2", dependencies = TRUE)
#install.packages("stringi", dependencies = TRUE)

library(stringi)
library(reshape2)
```

#Create ratings matrix with rows as users and columns as movies. We don't need timestamp
```{r}
ratingmat = dcast(ratings, userId~movieId, value.var ="rating",na.rm = FALSE)
```
#Removing users ids
```{r}
ratingmat = as.matrix(ratingmat[,-1])
```
#The recommendation package in R: recommenderlab. 
It provides a UBCF Model (User Based Collaborative Filtering)
Calculaing similarity among user ratings through the following methods: Jaccard similarity, Cosine similarity or Pearson similarity. 

#Using cosine similarity

```{r}
#Installing the package recommenderlab
install.packages("recommenderlab", dependencies= TRUE)
library(recommenderlab)
```

#Reducing size of ratings matrix for faster computation
before reducing size 45.8 Mb, it is called a "sparse matrix"
due to the large number of zeros, by removing them it is transform in a "dense matrix"

#Converting rating matrix to real rating matrix which makes it dense
```{r}
ratingmat = as(ratingmat, "realRatingMatrix")
```
Size of the matrix is now 1.7 Mb
Next step is normalize the matrix so recommendations come out unbiased.

#Normalize the ratings matrix
```{r}
ratingmat = normalize(ratingmat)
```
The recommender model of recommenderlab package is Recommender()

```{r}
#Reading about Recommender function
?Recommender
```

#Building the model
```{r}
#Create Recommender Model
#The parameters are UBCF and Cosine similarity.
# Taking 10 nearest neighbours

rec_mod = Recommender(ratingmat, method="UBCF", param = list(method = "Cosine", nn=10))

```

#Making predictions
starting with the first user 
```{r}
#Obtain top 5 recommendations for 1st user entry in dataset
Top_5_pred = predict(rec_mod, ratingmat[1], n=5)
```

To see the predictions of the model, convert them in a list and print them out:
```{r}
Top_5_List = as(Top_5_pred, "list")
Top_5_List

```
The results are in movieId format.Using the movies dataset, we can check the movie titles maping from movieId. 
```{r}
#installing package "dplyr" 
# Using R version 3.5.0 to be able to run package "plyr"
install.packages("dplyr")
library(dplyr)
```
```{r}
#Converting the list into a dataframe, changing column name to movieId
Top_5_df = data.frame(Top_5_List)
colnames(Top_5_df) = "movieId"

#since movieId is of the type integer in Movies data
#Typecasting id in the recommendations as well
Top_5_df$movieId = as.numeric(levels(Top_5_df$movieId))

#Merge the movie ids with names to get titles and genres
#install.packages("join", dependencies= TRUE)
names = left_join(Top_5_df, movies, by= "movieId")

#Print the titles and genres
names

```
Conclusion: considering similarity between users, the model recommends 
the movies listed above, where it can be seen the year of the movie released, and the movie genres.  
